\subsection{Generalized Linear Model} 

Generalized Linear Models are an extension of linear regression that allow for the response variable to have a non-normal distribution. Having a binary response variable, the relationship between the predictors and the response variable is not linear, so the logit link function,$$logit(p) =log\frac{p}{1-p}$$ is used to provide a transformation to ensure that the response is constrained between 0 and 1.

```{r glm, echo = FALSE}

# Create a data frame to hold the results
model_results <- data.frame(Model = character()
                             ,Data = character()
                             ,Accuracy = numeric()
                             ,Kappa = numeric()
                             ,F1 = numeric()
                             ,ROC = numeric()
                             ,Sensitivity = numeric()
                             ,Specificity = numeric())
```

```{r glm_models, echo = FALSE}
###############################################################################
# Generalized Linear Models
###############################################################################

# Enable parallel processing and reserve resources
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)

# Set up training conditions - must use LOOCV
fitControl <- trainControl(method = "LOOCV"
                           ,classProbs = TRUE 
                           ,summaryFunction = twoClassSummary
                           ,savePredictions = 'final')

# Full data set
set.seed(1234)
glm.model.ud <- caret::train(Label ~.
                          ,data = use_data
                          ,family = 'binomial'
                          ,method = 'glm'
                          ,trControl = fitControl
                          ,metric = "ROC")

# Get the confusion matrix
glm.cm.ud <- caret::confusionMatrix(glm.model.ud$pred$pred, glm.model.ud$pred$obs, mode = "everything")

# Get the performance metrics from the model and save for comparison
performance <- getTrainPerf(glm.model.ud)
model_results <- rbind(model_results
                        ,data.frame(Model = 'GLM'
                        ,Data = 'Full'
                        ,Accuracy = glm.cm.ud$overall[1]
                        ,Kappa = glm.cm.ud$overall[2]
                        ,F1 = glm.cm.ud$byClass[7]
                        ,ROC = performance[,1]
                        ,Sensitivity = performance[,2]
                        ,Specificity = performance[,3]))

# RFE selection data set
set.seed(1234)
glm.model.md <- caret::train(Label ~.
                          ,data = model_data
                          ,family = 'binomial'
                          ,method = 'glm'
                          ,trControl = fitControl
                          ,metric = "ROC")

# Get the confusion matrix
glm.cm.md <- caret::confusionMatrix(glm.model.md$pred$pred, glm.model.md$pred$obs, mode = "everything")

# Get the performance metrics from the model and save for comparison
performance <- getTrainPerf(glm.model.md)
model_results <- rbind(model_results
                        ,data.frame(Model = 'GLM'
                        ,Data = 'RFE Selection'
                        ,Accuracy = glm.cm.md$overall[1]
                        ,Kappa = glm.cm.md$overall[2]
                        ,F1 = glm.cm.md$byClass[7]
                        ,ROC = performance[,1]
                        ,Sensitivity = performance[,2]
                        ,Specificity = performance[,3]))

# Low correlation data set
set.seed(1234)
glm.model.lc <- caret::train(Label ~.
                          ,data = use_data_lc
                          ,family = 'binomial'
                          ,method = 'glm'
                          ,trControl = fitControl
                          ,metric = "ROC")

# Get the confusion matrix
glm.cm.lc <- caret::confusionMatrix(glm.model.lc$pred$pred, glm.model.lc$pred$obs, mode = "everything")

# Get the performance metrics from the model and save for comparison
performance <- getTrainPerf(glm.model.lc)
model_results <- rbind(model_results
                        ,data.frame(Model = 'GLM'
                        ,Data = 'Low Correlation'
                        ,Accuracy = glm.cm.lc$overall[1]
                        ,Kappa = glm.cm.lc$overall[2]
                        ,F1 = glm.cm.lc$byClass[7]
                        ,ROC = performance[,1]
                        ,Sensitivity = performance[,2]
                        ,Specificity = performance[,3]))

rownames(model_results) <- NULL

# Disable parallel processing and release resources
stopCluster(cluster)
registerDoSEQ()

```

```{r glm_roc, echo = FALSE}

kable(model_results, caption = "Model Results", booktabs = TRUE) %>% 
        kable_styling(font_size = 8, latex_options = c("striped", "hold_position"))

glm.ROC <- roc(glm.model.ud$pred$obs, glm.model.ud$pred$Normal)
plot(glm.ROC, col = 'blue', main = paste('GLM - Area under the curve (AUC):', round(auc(glm.ROC),2)))

# Print model coefficients
coefficients <- as.data.frame(glm.model.ud$finalModel$coefficients)

kable(coefficients, caption = "GLM Model Coefficients", booktabs = TRUE) %>% 
        kable_styling(font_size = 8, latex_options = c("striped", "hold_position"))

```

