# Exploratory Data Analysis

Note in __Table 1__ that the count of observations by segment differs. This indicates that not all of the `r length(unique(full$PatientNumMasked))` unique patients have observations for all lung segments. 

```{r observations, echo = FALSE}

observations <- use_data %>% 
        group_by(Position) %>%
        summarize(Count = n())

observations$Position <- as.character(observations$Position)
observations <- rbind(observations, c('Total', sum(observations$Count)))
kable(observations, caption = "Lung Segment Observations", booktabs = TRUE) %>% 
        kable_styling(font_size = 8,  latex_options = c("striped"))

```

Talk about checking to make sure labels for all patient ids are the same across tabs.

```{r check_labels, echo = FALSE}

# Verify that all labels for patients are the same
checkLabel <- full %>%
        group_by(PatientNumMasked) %>%
        mutate(minLabel = min(Label)) %>%
        mutate(maxLabel = max(Label)) %>%
        mutate(count = length(PatientNumMasked)) %>%
        select(PatientNumMasked, minLabel, maxLabel, count) 

```

The correlation plot in __Figure 2__ shows that some of the predictors have strong correlations with each other. This requires further investigation as high correlation among predictors can cause instability in the model. Generally, when two variables are highly correlated, it is best to remove one since the second is not providing new information.

```{r correlation, echo = FALSE, out.width = '350px', fig.cap = 'Correlation Matrix', fig.align = "center", fig.pos = 'H'}

# Correlation between predictors
corr <- cor(use_data[-c(40,41)])
corrplot(corr, type = "upper", tl.cex = 0.5, tl.col = "blue", tl.srt = 45)

```

```{r high_correlation, echo = FALSE}

# Review and remove highly correlated predictors
correlationMatrix <- cor(use_data[-c(40,41)])
highCorrelation <- findCorrelation(correlationMatrix, cutoff = 0.50)
rows <- length(colnames(use_data[highCorrelation]))
highly_correlated <- data.frame("Correlated Variables 1" = colnames(use_data[highCorrelation])[1:(rows / 2)], 
                                "Correlated Variables 2" = colnames(use_data[highCorrelation])[(rows/2 + 1):rows])

kable(highly_correlated, caption = "Highly Correlated Variables", booktabs = TRUE, col.names = c('', '')) %>% 
        kable_styling(font_size = 8, position = 'float_right', latex_options = c("striped", "hold_position"))

```

```{r new_correlation, echo = FALSE, out.width = '350px', fig.cap = 'Correlation Matrix - Highly Correlated Variables Removed', fig.align = "center", fig.pos = 'H'}
use_data_lc <- use_data[,-highCorrelation]
low_corr <- use_data_lc %>%
        select(-Position, -Label2)
low_corr<- cor(low_corr)

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(low_corr, method = "color", col = col(200),
         type = "upper", order = "hclust", number.cex = .7,
         addCoef.col = "black", # Add coefficient of correlation
         tl.col = "blue", tl.srt = 45, tl.cex = 0.5, # Text label color and rotation
         # Combine with significance
         sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag = FALSE)

```
